### 2016-12-20 tafaseer_group/03_results/appendix.rmd
Moving things around wildly. Sorry, too tired for detailled breakdown.

# Appendix

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff=60, indent=2), tidy=TRUE)
```

## I Script [`01_retrieval/download.r`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/01_retrieval/download.r)

```{r, code=readLines('../01_retrieval/download.r'), eval=FALSE}
```

## II Script [`01_retrieval/extract.r`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/01_retrieval/extract.r)

```{r, code=readLines('../01_retrieval/extract.r'), eval=FALSE}
```

## III Script [`01_retrieval/process.rb`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/01_retrieval/process.rb)

```{ruby, code=readLines('../01_retrieval/process.rb'), eval=FALSE}
```

## IV Script [`01_retrieval/convert.rb`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/01_retrieval/convert.rb)

```{ruby, code=readLines('../01_retrieval/convert.rb'), eval=FALSE}
```

## V Script [`02_analysis/wordcount_by_author.rb`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/02_analysis/wordcount_by_author.rb)

```{ruby, code=readLines('../02_analysis/wordcount_by_author.rb'), eval=FALSE}
```

## VI Query [`02_analysis/wordcount_whole_corpus.sql`](https://github.com/Islamicate-DH/hw/tree/master/tafaseer_group/02_analysis/wordcount_whole_corpus.sql)

```{sql, code=readLines('../02_analysis/wordcount_whole_corpus.sql'), eval=FALSE}
```